RCloud: DevOps for Visual Analysis and Data Science

### 2014-03-21 system.tex
yes

The internal computing infrastructure of organizations has changed
radically in the last fifteen years or so. Much of the move toward the
``cloud'' brought with it an ecosystem of processes distributed over a
network (many times virtually defined) that communicate over a
high-level protocol of some sort. Messages are typically exchanged
over HTTP. HTTP is attractive because web browsers and servers are
ubiquituous and available in essentially all pieces hardware today,
from tiny sensors in sensor networks to handheld devices to laptops to
rack-mounted servers.

As a result, in our view HTTP has become the lingua franca of IPC in
an intranet. One of the design goals for RCloud was to provide an
attractive environment for creators of data-analysis scripts, but at
the same time to create a system that behaves as a first-class citizen
in the ecosystem of computer services and networks with an
organization.

@@

We designed RCloud around the front-facing API, which has roughly one
entry point for each requirement.

@@

In RCloud, reputation and interest are a relationship between
\emph{notebooks} and \emph{users}, rather than a relationship between
user pairs. We decided on this approach because we expect typical
RCloud deployments to have relatively few users, but each user to have
 written relatively many notebooks.
In this case, assigning interest to users would not provide
``high-resolution'' data.

Explicit signaling of interest: starring.

Implicit signaling of interest: click-through count and execution
count.

These will be ingredients for the recommendation system.

@@

Notebooks by default are visible by the entire organization: there
exists a URL for every notebook in RCloud. This is deliberate. As
pointed out by Wattenberg and Kriss~\cite{Wattenberg:2011:DFS}, broad
access to analysis outputs (in their case, in the form of NameVoyager)
increases long-term engagement partly by the crossreferences in the
web. Our main RCloud deployment is only visible inside an intranet,
but we have nevertheless found anecdotal support for this theory by
noticing links to RCloud notebooks in internal discussion fora and
mailing lists.

@@

>>>>

The internal computing infrastructure of organizations has changed
radically in the last fifteen years. The shift from large
servers toward scalable, lower-cost, distributed systems (``the cloud'')
led to a software ecosystem of processes distributed over a
network (often virtually defined) communicating via some
high-level protocol, usually HTTP.  HTTP is dominant because web
browsers and servers are ubiquituous and available in essentially
all hardware devices, from tiny sensors, to handheld devices and
laptops, to rack-mounted servers.

As a result, HTTP is the lingua franca of interprocess communication
(IPC). One of the design goals for RCloud was to provide an attractive
environment for creators of data-analysis scripts, that also behaves
as a first-class citizen in the pre-existing ecosystem of computer
services and networks with an organization.

@@

We designed RCloud around the front-facing API, which provides roughly
one entry point to correspond with each requirement.

@@

In RCloud, reputation and interest are a relationship between
\emph{notebooks} and \emph{users}, rather than a relationship between
user pairs. We decided on this approach because we expect typical
RCloud deployments to have relatively few users, but each user to have
written relatively many notebooks. Under that assumption, assigning
interest to users would not provide sufficiently ``high-resolution'' data.

We incorporate both explicit and implicit indications of interest
in notebooks. Explicit interest is signalled by ``starring,'' or
clicking on a button that marks a notebook as interesting. 
This makes explicit indication of interest a nearly trivial operation,
always readily available, thus encouraging its use.

Implicit signaling of interest is supported by keeping click-through
\cite{Joachims:2005:AIC} and execution counts. (In addition to these
classic web search techniques of collecting feedback, we have the
prospect of applying static and dynamic code analysis to generate
further fine-grained information about relationships, for example,
which packages and data sets often appear together.)

\todo{explain} These will be ingredients for the recommendation system.

@@

There exists a URL for every notebook in RCloud, and notebooks by
default are visible by the entire organization. This is deliberate.
As pointed out by Wattenberg and Kriss~\cite{Wattenberg:2011:DFS},
broad access to analysis outputs (in their case, for NameVoyager)
increases long-term engagement in part through crossreferences on
the web. Although our prototype RCloud deployment is only visible
inside a corporate intranet, we nevertheless found anecdotal support
for this notion by discovering links to RCloud notebooks in internal
discussion fora and mailing lists.

@@

A key engineering decision in RCloud was to rely on full two-way
communication between human interface in a web client, and remote
data and computing resources in the cloud.
